### 2014-01-08 image_analysis_paper/references_for_plos.tex
Added bacground references readme file for Yannis

\documentclass[12pt]{article}

\title{References for PloS paper}

\usepackage{biblatex}

\bibliography{../../references/references}

\begin{document}

\section{Algorithms}

\fullcite{flusser2000independence}
\begin{itemize}
\item Construct moments (polynomials) of the pixel intensities that are invariant to rotations
\end{itemize}
%
\fullcite{besl1992method}
\begin{itemize}
	\item Iterative closest point (ICP) algorithm
	\item Aligns/registers point clouds
\end{itemize}
%
\fullcite{lowe1999object}
\begin{itemize}
	\item Scale-invartiant feature transform (SIFT)
	\item Computes feature vectors for image; features are scale, translation, and rotation invariant
\end{itemize}
%
\fullcite{sadler1992shift}
\begin{itemize}
	\item Calculate bispectrum for image
	\item Bispectrum is rotation-invariant
	\item Do not need to align the images
\end{itemize}
%
\fullcite{hilai1994recognition}
\begin{itemize}
	\item Construct rotation-invariant basis for images
\end{itemize}
%
\fullcite{ojala2002multiresolution}
\begin{itemize}
	\item Decribe local patches of image (texture) by binary code
	\item Define function on binary code that is rotation-invariant
	\item Classify using binary code
\end{itemize}
%
\fullcite{zhao2013fourier}
\begin{itemize}
	\item Construct PCA basis that is rotation-invariant
\end{itemize}

\section{Applications}

\fullcite{hajnal2010medical}
\begin{itemize}
	\item Medical image registration (MRI, CT, etc.)
\end{itemize}
%
\fullcite{greenspan1994rotation}
\begin{itemize}
\item Texture recognition/classification
\item use Fourier basis; fit texture response (in Fourier basis) to rotations 
\end{itemize}
%
\fullcite{rowley1998rotation}
\begin{itemize}
	\item face recognition
	\item train neural net to estimate face orientation, then orient faces
\end{itemize}
%
\fullcite{simard1992efficient}
\begin{itemize}
	\item recognize digits
	\item only consider small rotations; fit plane to span of rotations of each digit, then see which plane new digit is closest to
\end{itemize}

\end{document}