### 2014-01-29 Amit_notes/Procrustes.tex
Updated notes from Amit

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage{amssymb}

\title{Solving the Procrustes Problem with PCA}

\begin{document}

\maketitle

Assume we have a signal that consists of $n$ features, each of which lie in $d$-dimensional space (e.g. if we have a 10-atom molecule in three-dimensional space, then $n=10$ and $d=3$).
%
Denote the features as $x_1, \dots, x_n \in \mathbb{R}^d$.

We observe $m$ rotated noisy copies of the features
\begin{equation}
y_i^{(k)} = O_k x_i + \xi_i^{(k)}, k=1, \dots, m, i=1, \dots, n,
\end{equation}
where $O_k$ is an orthoganol transform, and $\xi_i^{(k)}$ is Gaussian white noise. 
%
We would like to estimate the original signal $x_1, \dots, x_n$ from the rotated noisy copies $y_i^{(k)}$.

Note that, because orthoganol transforms preserve inner products,
\begin{equation}
\langle y_i^{(k)}, y_j^{(k)} \rangle  \approx \langle x_i, x_j \rangle 
\end{equation}
(up to noise).

Therefore,  we can estimate $\langle x_i, x_j \rangle$ as 
\begin{equation} \label{eq:ip_est}
\langle x_i, x_j \rangle  \approx \frac{1}{m} \sum_{k=1}^m \langle y_i^{(k)}, y_j^{(k)} \rangle 
\end{equation}

We then construct the matrix $C$, where $C_{ij} = \langle x_i, x_j \rangle$, using the estimates in \eqref{eq:ip_est}.
%
However, $C$ is precisely the covariance matrix of the data $x_1, \dots, x_n$. 
%
We can therefore recover $x_1, \dots, x_n$ using eigendecomposition:
\begin{equation}
\begin{bmatrix}
x_1^T \\
\vdots \\
x_n^T 
\end{bmatrix} = 
\begin{bmatrix}
| & | &  & | \\
\phi_1 & \phi_2 & \cdots & \phi_d \\
| & | &  & | \\
\end{bmatrix}
\begin{bmatrix}
\sqrt{\lambda_1} & 0 & 0 & \cdots & 0 \\
0 & \sqrt{\lambda_2} & 0 & \cdots & 0 \\
\vdots & \vdots & \vdots & \ddots & \vdots \\
0 & 0 & 0 & \cdots & \sqrt{\lambda_d} 
\end{bmatrix}
\end{equation}
where $\lambda_1, \dots, \lambda_n$ and $\phi_1, \dots, \phi_n$ are the eigenvalues and eigenvectors, respectively, of $C$, and they are ordered such that $\lambda_1 \ge \lambda_2 \ge \dots \ge \lambda_n$.

\end{document}