### 2014-02-13 image_analysis_paper/paper_draft.tex

\documentclass[12pt]{article}

\begin{document}

\section{Introduction}

% only images

\section{Techniques}

% show schematic with images (starfish or zebrafish w/ stripes)

\subsection{Alignment Using Angular Synchronization}

\subsection{Ordering Using Diffusion Maps}

\section{Example: {\em Drosophila}}

\subsection{dpERK}

\subsection{Dorsal}

\section{Discussion}
% compare to TSP algorithms

\end{document}