### 2014-02-14 image_analysis_paper/background_reading/article_notes.tex
added edits to paper, backgrond reading; added eve data

\documentclass[12pt]{article}

\usepackage{biblatex}

\bibliography{references}

\begin{document}

\begin{enumerate}

\item \fullcite{wei2006conditional}

\item \fullcite{wei2006quantile}

\item \fullcite{li2009stochastic}

\item \fullcite{jordan2013behavioral}

\item \fullcite{tucker2010pseudotemporal}
\begin{itemize}
	\item Uses cross-sectional data of eye disease progression (e.g. glaucoma).
	\item Build time series from cross-sectional data (``pseudo time series'')
	\item Subsample data; from each subsample, build a ``pseudo time series'' by finding the shortest path through the points
	\item Use these shortest paths as time series to train a time series model (i.e. a HMM model).
\end{itemize}

\item \fullcite{kafri2013dynamics}
\begin{itemize}
	\item Take snapshot of fixed cell population
	\item Fit 1D curve to trajectory (they assert that the two ``important'' parameters are levels of DNA and the level of ``anaphase-promoting complex'', so they project onto these two components and then fit a 1D closed curve to the data)
	\item Since the cells follow a closed cycle, given the closed curve, they can use mass balances to determine the relative rates/fluxes through the different stages
\end{itemize}

\item \fullcite{wyart2010evaluating}

\begin{itemize}
	\item Use RNA FISH technique to count levels of individual mRNAs in cells; requires cells to be fixed
	\item Propose models for dynamics of mRNA throughout cell cycle
	\item Show how they can use data from population of fixed cells to fit model (periodic model for level of mRNA production)
\end{itemize}

\item \fullcite{chen2012nonlinear}

\item \fullcite{das2013dynamic}

\item \fullcite{schindler2007inferring}

\item \fullcite{bigot2012geometric}

\item \fullcite{escudero2011epithelial}

\item \fullcite{machta2013parameter}

\item \fullcite{peng2013detecting}

\item \fullcite{gerber2010visual}

\item \fullcite{qiu2011discovering}

\item \fullcite{gupta2008extracting}

\item \fullcite{lau2009modeling}

\item \fullcite{toga2003temporal}

\item \fullcite{li2012modelling}

\item \fullcite{denkinger2013set}

\item \fullcite{xu2012comprehensive}

\item \fullcite{sisan2010event}

\item \fullcite{golparvar2011monitoring}

\item \fullcite{golani1999phenotyping}
\begin{itemize}
	\item Want to construct video of construction progress from snapshots
	\item Building Information Models (BIM) 
\end{itemize}

\item \fullcite{anavy2014blind}

\end{enumerate}

\end{document}