### 2014-05-21 image_analysis_paper/fig5.tex
edits prior to Yannis on 5/21

\documentclass[12pt]{minimal}

\usepackage{graphicx}
\graphicspath{ {paper_figures2/} }

\usepackage{tikz}

\usepackage[margin=0in, paperwidth=8.7cm, paperheight=9.5cm]{geometry}

\begin{document}

\centering
%\raisebox{5.4cm}{\textsf{A}}
\begin{tikzpicture}
     \node at (0,0)
       {\includegraphics[width=8cm]{mut_wt_vdm_embedding_background}};
     \node at (0,0)
       {\includegraphics[width=8cm]{mut_wt_vdm_embedding}};
      \node at (-4cm, 2.4cm) {\textsf{A}};
\end{tikzpicture}\\
\raisebox{0.6cm}{\textsf{B}}
\includegraphics[width=8cm]{wt_trajectory}\\
\raisebox{0.6cm}{\textsf{C}}
\includegraphics[width=8cm]{mut_trajectory}

\end{document}