### 2014-05-26 image_analysis_paper/fig5.tex
updated mutant figures and some text for paepr

\documentclass[12pt]{minimal}

\usepackage{graphicx}
\graphicspath{ {paper_figures2/} }

\usepackage{tikz}

\usepackage[margin=0in, paperwidth=8.7cm, paperheight=9.5cm]{geometry}

\begin{document}

\centering
%\raisebox{5.4cm}{\textsf{A}}
\begin{tikzpicture}
     \node at (0,0)
       {\includegraphics[width=8cm]{mut_wt_vdm_embedding_backgroundb}};
     \node at (0,0)
       {\includegraphics[width=8cm]{mut_wt_vdm_embeddingb}};
      \node at (-4cm, 2.4cm) {\textsf{A}};
\end{tikzpicture}\\
\raisebox{0.6cm}{\textsf{B}}
\includegraphics[width=8cm]{wt_trajectoryb}\\
\raisebox{0.6cm}{\textsf{C}}
\includegraphics[width=8cm]{mut_trajectoryb}

\end{document}