### 2014-05-29 image_analysis_paper/paper_figures2/spheres.tex

\documentclass{minimal}

\usepackage{tikz}           %for TikZ graphics
\usepackage{tikz-3dplot} %for 3dplot functionality

%\usepackage[active,tightpage]{preview}      %generates a tightly fitting border around the work
%\PreviewEnvironment{tikzpicture}
%\setlength\PreviewBorder{2mm}

\def\R{1}

\begin{document}
\tdplotsetmaincoords{70}{135}
\tdplotsetrotatedcoords{45}{0}{0}
\begin{tikzpicture}[line join=bevel,tdplot_rotated_coords, fill opacity=.7]

\tdplotsphericalsurfaceplot{36}{18}%
{\R}{black}{red}%
    {\draw[color=black,thick,<->] (-2,0,0) -- (2,0,0) node[anchor=north east]{$y$};}%
    {\draw[color=black,thick,<->] (0,-2,0) -- (0,2,0) node[anchor=north west]{$x$};}%
    {\draw[color=black,thick,<->] (0,0,-2) -- (0,0,2) node[anchor=south]{$z$};}%

    \tdplotsetpolarplotrange{70}{120}{60}{120}
\tdplotsphericalsurfaceplot{36}{18}%
{\R}{black}{blue}%
{}{}{}%

\end{tikzpicture}

\end{document}