### 2015-02-11 image_analysis_paper_development2/paper_figures/fig5.tex

\documentclass[10pt]{article}

\usepackage[paperwidth=18cm, paperheight=4.5cm, margin=0cm]{geometry}
\usepackage{graphicx}
\hyphenpenalty=10000

\begin{document}

\begin{minipage}[b]{12cm}
\raisebox{0.5cm}{\textsf{A}}
\includegraphics[width=12.25cm]{drosophila_fixed_images_scrambled}

\raisebox{0.5cm}{\textsf{B}}
\includegraphics[width=12.25cm]{drosophila_fixed_images_ordered}

\raisebox{0.5cm}{\textsf{C}}
\includegraphics[width=12.25cm]{drosophila_fixed_images_average}
\end{minipage}
%
\hfill
%
\raisebox{3.25cm}{\textsf{D}}
\includegraphics[width=4.25cm]{drosophila_fixed_images_rank_corr}

\end{document}